Response to Reviewers

### 2021-11-09 Scripts/Manuscript.Rmd
Addressed most comments

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods). **We define a ‘stock’ unit as a species shared by neighboring EEZs.** Firthermore, we  developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1N)). By employing the concept of *time of emergence*, we estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig. 1). Such estimation is based on two thresholds representing a 67 and a 95% probability of shift. Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods). This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).

```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.pos="H", fig.cap =  "Schematic diagram of the transboundary index (TI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another. The index is based on the distance between the distributional centroid of the transboundary stock and the geographic centroid of the neighboring EEZ sharing the stock (Top panel). Time of emergence is defined as the first year when the TI overshoots historical values (Bottom panel). Only showing a subset of ensemble members for clarity."}

@@

<!-- We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1)A). We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution. The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence". We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 67% that the index has emerged). -->

@@

We developed a Transboundary Index (TI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change. This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig. \@ref(fig:toeFigure1)). The centroid of a transboundary stock was determined by the average ($\mu$) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock. Therefore,

@@

where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes. Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TI) as follows:

\begin{equation}
\label{eq:six}
TI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
\end{equation}

where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$. The TI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.

<!-- Removed in response to reviewer 3 -->
<!-- Thus, a higher TI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels. From the perspective of a state’s fisheries, increasing TI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries. In contrast, a decreasing TI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries. -->

@@

Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers by when, ideally, climate adaptation plans will need to have been implemented [@Link:2010ei]. To that end, we adopted the concept of time of emergence, commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev] and defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (i.e., historical natural variation) [@Hawkins:2012hc]. The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc]. The TI’s natural internal variability was estimated from the ensemble members by firstly averaging the TI from 1951 to 2005 and then estimating the variation ($TI_\sigma$) across ensemble members. The TI signal represents the average TI across ensemble members from 2006 to 2100 ($TI_\mu$). We then set a threshold (*tresh*) of $TI_\sigma$ to define the time of emergence of transboundary stocks (See Fig. \@ref(fig:toeFigure1) for a graphical description):

<!-- We then set a threshold of two times $TI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig. \@ref(fig:toeFigure1) for a graphical description): -->

<!-- $EmergenceTIndex = \frac{\mu_{TI[t]}}{SD_{TI_{th}}}$ -->
<!-- $ToE = t when Emergence index[t] > 1$ -->

\begin{equation}
\label{eq:seven}
ToE = TI_\mu[t']> tresh
\end{equation}

This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TI “emerges” from historical internal variability (*tresh*). It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution. Two different arbitrary values of *tresh* were used. A conservative value ($tresh = 2* TI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TI_\sigma$) for a 67% probability.

@@

Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig. \@ref(fig:toeFigure1)A), using a two s.d. threshold (i.e., representing a probability of 95% that the stock has shifted). This corresponds to 45% of the studied stocks (Fig. \@ref(fig:toeFigure1)C). Using a less conservative threshold (i.e., one s.d. representing a probability of 67% that the stock has shifted) results in 33% (n = 5,745) more shifting stocks, that is, 67% of all studied stocks (Figs. \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)). For both cases, the first shift is modelled to have occurred back in 2006. The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (one s.d. threshold) and 2036 $\pm$ 28 years (two s.d. threshold). Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stock shift by the UN 2030 Agenda’s deadline (Fig. \@ref(fig:toeFigure1)C).

@@

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons display average year of range shifts within them. Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colors represent a late shift/low fishing revenue. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue. B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year of range shift. Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand."}

>>>>

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods). **We define a ‘stock’ unit as a species shared by neighboring EEZs.** Firthermore, we  developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1N)). By employing the concept of *time of emergence*, we estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig. 1). Such estimation is based on two thresholds representing a 67 and a 95% probability of shift. Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (Methods). This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).

```{r toeFigure1N, eval = T,  echo = F, fig.width=14, fig.height=10, fig.pos="H", fig.cap =  "Schematic diagram of the transboundary index (TSI) to determine the time of emergence of transboundary stocks from one Exclusive Economic Zone (EEZ) to another. **Hypothetical distribution range and shape of a shared stock.** The index is based on the distance between the distributional centroid of the transboundary stock and the centroid of the neighboring EEZ sharing the stock (Top panel). Time of emergence is defined as the first year when the TSI overshoots historical values (Bottom panel). Only showing a subset of ensemble members for clarity."}

@@

<!-- We then developed a Transboundary Shift Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1)A). We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution. The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence". We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 67% that the index has emerged). -->

@@

We developed a Transboundary Shift Index (TSI) to evaluate range shifts in the shared distribution of transboundary stocks under climate change. This index represents the shift in the distribution centroid of a transboundary stock relative to the centroid of neighboring EEZs that share this stock (Fig. \@ref(fig:toeFigure1)). The centroid of a transboundary stock was determined by the average ($\mu$) latitude ($lat_{ts}$) and longitude ($lon_{ts}$) across grid cells within the neighboring EEZs sharing the stock. Therefore,

@@

where $lat_{eez}$ and $lat_{ts}$ are the latitudes of the EEZ and transboundary stock centroids, respectively, and $lon_{eez}$ and $lon_{ts}$ are associated longitudes. Then, for each year and for each ensemble member between 1951 and 2100 we calculated the transboundary index (TSI) as follows:

\begin{equation}
\label{eq:six}
TSI =(\frac{D_{A,t}}{sd(D_{A,th})} - \frac{D_{B,t}}{sd(D_{B,th)}})^2
\end{equation}

where $D_A$ and $D_B$ represent the distance between the distribution centroids of a stock and the centroid of the neighboring EEZs (A and B) sharing the stock for each time step from 2006 to 2100 (*t*); and *sd* is the standard deviation of the historical (th, 1951 - 2005) centroid distribution for $D_A$ and $D_B$. The TSI was smoothed to a 10-year average to reduce interannual variability [@Frolicher:2016dg] and to match transboundary fisheries management that often integrates a longer time-frame in their consideration.

<!-- Removed in response to reviewer 3 -->
<!-- Thus, a higher TSI means that the abundance of a transboundary stock within neighboring EEZs changes beyond historical levels. From the perspective of a state’s fisheries, increasing TSI indicates an increase in the challenges to effective management and sustainability of the transboundary stock’s fisheries. In contrast, a decreasing TSI suggests that the sharing of the stocks between neighboring EEZs is becoming relatively more stable; a condition that favors effective management and sustainability of the related fisheries. -->

@@

Knowing the point in time at which the distribution of a shared stock will diverge from historical variability is important to inform decision makers by when, ideally, climate adaptation plans will need to have been implemented [@Link:2010ei]. To that end, we adopted the concept of time of emergence, commonly applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev] and defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (i.e., historical natural variation) [@Hawkins:2012hc]. The premise behind time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc]. The TSI’s natural internal variability was estimated from the ensemble members by firstly averaging the TSI from 1951 to 2005 and then estimating the variation ($TSI_\sigma$) across ensemble members. The TSI signal represents the average TSI across ensemble members from 2006 to 2100 ($TSI_\mu$). We then set a threshold (*tresh*) of $TSI_\sigma$ to define the time of emergence of transboundary stocks (See Fig. \@ref(fig:toeFigure1) for a graphical description):

<!-- We then set a threshold of two times $TSI_\sigma$ representing a probability of 95% that the transboundary index has “emerged” from historical variability (See Fig. \@ref(fig:toeFigure1) for a graphical description): -->

<!-- $EmergenceTSIndex = \frac{\mu_{TSI[t]}}{SD_{TSI_{th}}}$ -->
<!-- $ToE = t when Emergence index[t] > 1$ -->

\begin{equation}
\label{eq:seven}
ToE = TSI_\mu[t']> tresh
\end{equation}

This way we define a shift in a stock’s shared distribution (i.e., time of emergence) as the first year the TSI “emerges” from historical internal variability (*tresh*). It assumes that, from a fisheries’ perspective, the first year a stock’s distribution shift from a 10-year average is sufficient to call for caution. Two different arbitrary values of *tresh* were used. A conservative value ($tresh = 2* TSI_\sigma$) representing a 95% probability that the index has emerged from historical variability and a more relaxed value ($tresh=TSI_\sigma$) for a 67% probability.

@@

Our results suggest that 4,119 transboundary stocks will experience a range shift beyond historical variability by 2100 (hereafter referred to as ‘shifts’; Fig. \@ref(fig:toeFigure1)A), using a two s.d. threshold (i.e., representing a probability of 95% that the stock has shifted). This corresponds to 45% of the studied stocks (Fig. \@ref(fig:toeFigure1)C). Using a less conservative threshold (i.e., one s.d. representing a probability of 67% that the stock has shifted) **results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks** (Figs. \@ref(fig:toeFigureS2) and \@ref(fig:toeFigureS8)). For both cases, the first shift is modelled to have occurred back in 2006. The average year of shift in the shared distribution of these stocks in all EEZs analyzed is projected to be between 2029 $\pm$ 27 years (one s.d. threshold) and 2036 $\pm$ 28 years (two s.d. threshold). Furthermore, between 81% and 83% of the world’s EEZs will experience at least one transboundary stock shift by 2100 with 77% to 80% of EEZs recording a distributional stock shift by the UN 2030 Agenda’s deadline (Fig. \@ref(fig:toeFigure1)C).

@@

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Year of change in the shared distribution of 4,119 transboundary stocks. A) Land polygons show the contribution of shifting stocks to a country or territory’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygons display average year of range shifts within them. Warm colors are indicative of an early shift/high fishing revenue contribution from transboundary stocks while cool colors represent a late shift/low fishing revenue. EEZs with no distributional shift between 2006 and 2100 are represented in pale blue. B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region. Horizontal dashed line represents the year by which countries have committed to reach full implementation of the 2030 Agenda. N = North, S = South, W = West and E = East. Ltn. Ame. and the Car. = Latin America and the Caribbean. Aus and New Z. = Australia and New Zealand."}

### 2021-11-09 Scripts/responses_rev_gcb.Rmd
Addressed most comments

@@

@@

@@

@@

@@

- **Response:**
Fig. 1: This is a wonderful schematic, my concerns: 1) stocks that do not occur as ovals but occur in patches or blobs instead-- the centroid may be too crude to detect movement of non-centric distributions- I don;t think your approach should change but re-cast as a proof of concept, researchers interested in applying the technique would be able to apply it at a regional level, and perhaps adapt it to patch distributions (separate centroids).
2) I don;t know what the empty red circle means in bottom panel, I can’t find it. do you mean the white area since the red dot?
3) what is diff btwn distributional and geographic centroid? (suggest using one term or other).
4) WHat if EEZ Ca moves at the same rate in the same direction as EEZ Cb. Is the TI only useful if one (more northerly) stock is static? What about adding and index of each Ca and EEZb distance to border? Given how great this MS is, I imagine a lot of thought went through what constitutes a “shift”, so If I have misunderstood index. OH! Now I think I get it, after reading equations 5 and 6, is it possible to include more details in Fig 1?

@@

@@

>>>>

- **Response:** Thank you for pointing this out. In the original manuscript we were showing the percentage difference between the two numbers (i.e., $x = \frac{|45-63|}{|\frac{45+63}{2}|}*100$), hence the 33.3% difference. However, now we realize this can be misleading and have adopted a more straightforward computation. We also corrected the 63% typo (not 67%). It now reads:

  - - Line XX; *Using a less conservative threshold (i.e., one s.d. representing a probability of 67% that the stock has shifted) results in 18% (n = 5,745) more shifting stocks, that is, 63% of all studied stocks*

@@

- **Response:** Thank you for your commnets on our fiugre 3. We have devided the responses by each point you make. Please see below

  While I like the idea of Figure 3, I feel that there is too much conveyed in this figure. It is confusing to have two different scales (for land polygons and EEZs, respectively) on the same color ramp, even though they show completely different things.

  - **Response:**  **NOTE: I do not (fully) agree with this comment. yes, the figure is complex, but I do not think having different color scales will make it any less complex. I actually think it will make it even harder to read. Do you think we can counter-argue this comment?**

  Moreover, see my comments on averaging above.

  - **Response:** Please see our response to the comment above.

  Why is average year used for coloring EEZs, but median year used in Fig 3B?

  - **Response:** In the case of Fig.3B we used the median because this figure talks directly to the statistical non-parametric analysis (Kruskal-Wallis) looking at differences in the time of emergence by geographic region. **NOTE: Shall we change Fig.3A to show the median instead of the mean, for the sake of consistency (and make the reviewer happy)?**

@@

- **Response:** Yes, this refers to the median emergence year for all shifting stocks in each region. This is now included in the figure legend.

  - - Fig. 3 legend, *B) Year of shared distribution shifts by UN sub-regions. Points color coded by the median year for all shifting stocks in the region.*

@@

- **Response:** We intended to show this information in Figure 4. However, the message in the figure is not clear, as the reviewer pointed out in the next comment. We have modified figure 4 to show this as a figure.

@@

- **Response:** Thank you so much for taking the time to revise our manuscript and for your words of encouragement.

@@

- **Response:** Thank you for this comment. We have now changed the term to "*Transboundary Shift Index (TSI)*"

Fig. 1: This is a wonderful schematic, my concerns: 1) stocks that do not occur as ovals but occur in patches or blobs instead-- the centroid may be too crude to detect movement of non-centric distributions- I don;t think your approach should change but re-cast as a proof of concept, researchers interested in applying the technique would be able to apply it at a regional level, and perhaps adapt it to patch distributions (separate centroids).

- **Response:** We agree with the reviewer that the distribution of fish stocks in the ocean is not necessarily oval. In fact, the distribution of some of our stocks does not follow this patter. Nevertheless, we believe our method responds well to other patterns (as long as they are considered one single stock unit) since the centroid of the stock's distribution is based on the "high biomass-level" grid cells. We have included a clarification regarding the stocks distribution in the figure legend of the revised manuscript.

  - - *Hypothetical distribution range and shape of a shared stock.*

2) I don;t know what the empty red circle means in bottom panel, I can’t find it. do you mean the white area since the red dot?

- **Response:** Red circles represent the years the TI is larger than the historical variation. These were quite faded in the original diagram and have made them clearer in the revised figure.

3) what is diff btwn distributional and geographic centroid? (suggest using one term or other).

- **Response:** The distributional centroid represent's the stocks centroid while "geographical" centroid is the EEZs centroid. We have changed the second term as suggested by the reviewer, it is now called simply "EEZ centroid".

4) WHat if EEZ Ca moves at the same rate in the same direction as EEZ Cb. Is the TI only useful if one (more northerly) stock is static? What about adding and index of each Ca and EEZb distance to border? Given how great this MS is, I imagine a lot of thought went through what constitutes a “shift”, so If I have misunderstood index. OH! Now I think I get it, after reading equations 5 and 6, is it possible to include more details in Fig 1?

- **Response:** Ca and Cb represent the centroid of the EEZs, and thus, they do not move. What moves is the centroid of the stock (the smaller blue circle) along the whole distribution of the stock that spans both EEZs. We realize this was not clear in the legend and have now change it.

  - - : *say EEZ $C_A$: Centroid of the Exclusive Economic Zone A and EEZ $C_B$: Centroid of the Exclusive Economic Zone B*

@@

- **Response:** See our previous response on SDG 17

@@

- **Response:** Thank you very much for your kind words
